### 2022-02-19 manuscript/fig_wflamba.tex
fill the structure

\begin{figure}[ht!]
    \begin{subfigure}{\linewidth}
        %\caption{}
        \centering
        % include first image
        \includegraphics[width=1.0\linewidth, trim=-0cm 2cm 0 0cm]{fig/wflambda/chr6_pan_fa_a2fb268_4030258_6a1ecc2_smooth_C4_bad_sorted}
        \label{fig:bad-sorting}
    \end{subfigure}
%	\includegraphics[width=\linewidth]{fig/metrics/chr4.pan.HTTex1.gfa.multiqc_odgi_stats.png}
    \caption{
        PLACEHOLDER IMAGE XXX XXX XXX
    }
    \label{fig:wflambda}
\end{figure}